Combination of field and experimental data together with computational models

### 2022-02-28 manuscript_1.0/manuscript_1.0.Rmd
New version of the manuscript. Using figure from the old data.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,warning = FALSE,message=FALSE)
library(here)
source(file=here("loadFieldData.R"))
```

Disentangling the mechanistic underpinnings of behaviour 
is typically done under controlled conditions in a laboratory setting. However, 
controlled laboratory conditions often mask the inter and intra-specific 
variation that arises in the interaction between mechanisms and 
environmental variation. From a evolutionary perspective, mechanisms 
are likely selected because of how they allow individuals to respond to 
environmental variation. For example, biological market theory predicts that 
the exchange rate of goods and/or services traded between cooperative 
partners adjusts to the law of supply and demand, when individuals have 
some degree of partner choice [@noe_Biological_1995a]. 
Supply and demand conditions, which typically 
depend on the abundance of the species involved, certainly vary in time 
and space. Therefore, natural selection should 
favour the ability to flexibly adjust decisions and behavioural output to 
current market conditions. Indeed, such adjustments have been repeatedly 
documented [@axen_Signalling_1996]. In animals, an obvious general candidate 
mechanism for the strategic adjustment is the cognitive machinery. However, 
it is not clear which cognitive mechanisms allow individuals to adjust their 
behaviour to the varying conditions. As adjustments to changes in 
market conditions have been documented in a great variety of species with 
highly variable brain anatomies, the question arises to what extent mechanisms 
beyond basic associative learning may be involved.

One example of strategic adjustment in a biological market is the marine
cleaning mutualism involving the cleaner fish *Labroides dimidiatus* and
'client' fish. Client fish seek cleaner fish services at their territory 
(so-called “cleaning station”) and offer themselves as food patches 
to get their ectoparasites removed, which provides cleaners 
with food and clients with improved health [@waldie_LongTerm_2011;
@ros_Does_2011;@triki_Effects_2016;@demaire_Reduced_2020]. 
Given the capacity of some client fish to swim larger distances and 
access multiple cleaning stations while others access the only cleaning 
station at their territory, it is crucial to categorize clients as 
either “visitors” or “residents”, respectively. During cleaning interactions, 
a cleaner fish often face a choice between a visitor and a resident client 
seeking its cleaning services simultaneously. Visitors have the option 
to switch to another cleaner fish if being made to wait, while residents 
must wait for inspection. Indeed,
visitors have been observed to use their partner choice option in that
way [@bshary_Choosy_2002], which may explain why cleaners give
visitors service priority in a field study in the Red Sea
[@bshary_Cleaner_2001a].

When researchers aimed at testing wild-caught cleaner fish, as 
well as individuals from other species, abilities 
to prioritize visitor over resident clients in a lab-based paradigm, 
they used Plexiglas plates of different colours and shapes offering same 
amount of food as surrogates for visitor and resident clients.
One plate is made to behave like a resident, i.e. it remains until 
the cleaner fish had eaten off it. The other plate was made to behave 
like a visitor, i.e. it is removed if not inspected first.
Cleaners learned to prefer the visitor plate and
hence obtained the double amount of food [@bshary_Asymmetric_2002].
Furthermore, adult cleaner fish outperformed various primates as well as rats and
pigeons in this original version of the biological market task
[@salwiczek_Adult_2012; @zentall_Early_2017]; yet the African grey
parrots solve this task as well [@pepperberg_Can_2014]. Given that primates 
are known to readily distinguish 
between one and two reward units if presented in more conventional ways 
$\color{red}{\text{ref}}$, this raises the question:
what makes the task difficult for most non-cleaners?

@@

To solve the ephemeral reward task, cleaners need to account for 
the future consequences of current decisions. In the model by 
Quiñones *et al.* [@quinones_Reinforcement_2019], this could be achieved in 
two non-mutually exclusive ways: through low temporal discounting of 
future effects, also termed ‘chaining’ [@enquist_Power_2016]; and/or 
through perceiving a visitor client leaving as psychological punishment 
(i.e. as a negative reinforcer). Low temporal discounting is when 
individuals include in their valuation of an action the reward effects 
that this will have in the future. This is done by combining in a single 
valuation the reward obtained in the current time with all the reward that 
comes after, discounting for how far in the future reward 
is accrued. 'Chaining' the reward of these different time steps allows 
individuals to take actions that increase the long-term reward at the sacrifice
of short term considerations [@enquist_Power_2016]. Even though, 'chaining' 
can be readily implemented computationally in learning models 
[@enquist_Power_2016;@sutton_Reinforcement_2018], cognitively it seems to be 
a complex adaptation [@suddendorf_Evolution_2007]. On the other hand,
using client behaviour as a negative reinforcer is in principle 
easier to implement. Negative reinforcement is part of the ubiquitous 
learning mechanism termed operant conditioning [@thorndike_Animal_1898; 
@skinner_Behavior_1938]. Thus, standard logic of 
Lloyd Morgan’s cannon demands that operant conditioning as the simpler 
explanation is to be accepted by default. Ideally, the two mechanisms should 
be evaluated in light of how well they explain the available data.

@@

The interplay between the underling cognitive machinery and 
local ecological conditions apparently generate the documented variable 
performance among individuals of the same species. 
Our approach of fitting the computational model to the empirical data on 
fish densities and cleaner fish censuses and performance
in the ephemeral reward task aimed at: (i), determining which 
mechanism cleaner fish use to incorporate future consequences of current 
decisions by testing whether operant conditioning, chaining, or a combination 
of both best explains their performance; (ii)  determining whether the 
two mechanisms differed with respect to the ecological conditions 
that are likely to cause high versus low performance in the 
ephemeral reward task. Additionally, we assessed which mechanism 
yields optimal performance patterns. Relying on the logic of 
biological market theory, we predicted 
that appropriate performance is to show low preference for visitors under low
local cleaner-to-client ratio. 
<!-- In this case, as visitor leaving rates  -->
<!-- should be low as well and any leaving visitor will be readily  -->
<!-- replaced by another client in the queue. In contrast,  -->
<!-- cleaners should show high performance in the ephemeral reward task -->
<!-- if the local cleaner-to-client ratio favours clients.  -->

@@

The model shows that agents need to find a mean 
to incorporate future consequences of current choices. In the model, 
this could be achieved with either of two parameters that could 
also work together.  First, $\gamma$ measures how
much individuals include future rewards in their decision updates. If
$\gamma=0$, individuals only use the immediate reward obtained from a
cleaning interaction. As $\gamma$ increases individuals include more the
reward obtained from the subsequent choices. That amounts to calculating
and using for decision making the future expected rewards of an action.
Second, $\eta$ measures how much individuals include in their reward the
fleeing behaviour of visitor as a negative component. Both of
these parameters allow individuals to use in their estimates the future
effects of their choices.

@@

<!-- Cleaner densities, in interaction with visitor client leaving probability,  -->
<!-- explained best individual performance in the ephemeral reward task  -->
<!-- [@triki_Biological_2019], while cleaner-to-client ratio was not a significant  -->
<!-- factor. This creates a certain mismatch with the model by Quiñones *et al.*  -->
<!-- [@quinones_Reinforcement_2019], which focused on abundance of cleaners relative  -->
<!-- to overall client abundance in order to capture the characteristics  -->
<!-- of the market. Importantly, relative abundances of clients define not only  -->
<!-- the probability of interactions with residents and visitors, but also affect  -->
<!-- how often the cleaning station is empty (e.g. there are no clients to be cleaned).  -->
<!-- The higher the relative abundance of cleaners is, the higher is the probability  -->
<!-- of cleaning stations being empty at a given time point. How often the  -->
<!-- cleaning stations are empty in turns influences the value that cleaners can  -->
<!-- expect from their choices. The emptier the cleaning stations are, the more  -->
<!-- beneficial it is to prioritize a client that will not stay. On the other hand,  -->
<!-- situations where cleaners have to choose will be rare if cleaner fidh relative  -->
<!-- abundance is high, which reduces the total value of prioritizing visitors  -->
<!-- over residents and makes it more difficult to learn a preference for visitors.  -->
<!-- Given the model structure, we used the empirical data to calculate relative  -->
<!-- abundances of cleaners, residents and visitors for the current analyses.  -->
<!-- To incorporate the empirical finding that absolute densities matter, we  -->
<!-- introduced a new scaling constant (see section on model fitting).  -->

@@

The cognitive performance data was from a total of 120 cleaners 
(10 individuals per 12 site/year) tested in the 
ephemeral reward task [@triki_Biological_2019;@triki_Brain_2020]. Authors 
housed all captured cleaners individually in glass aquaria 
( $62cm \times 27cm \times 37 cm$ ) and provided them 
with PVC pipes ($10 cm \times 1 cm$) as shelters. 
The task consisted of exposing the cleaner fish to substitute 
models of client fish in the form of two *Plexiglas* plates offering the 
same amount of food (one item of mashed prawn).  The two plates differed 
in colour and pattern (horizontal green stripes or vertical pink stripes) 
but had equal size ($10 cm \times 7 cm$). Importantly, the two plates played 
different roles as either a visitor (ephemeral food source) or 
resident (permanent food source). That is, if a cleaner fish inspected the 
resident plate first, the experimenter withdrew the visitor plate out of 
the aquarium as a consequence.  Choosing first the visitor plate, 
however, granted access to both plates. The equal size of the plates 
forced cleaner fish to learn to give service priority to the visitor plate 
based solely on the behaviour-cue of the plates rather than size-cue 
[@wismer_Cuebased_2019]. Triki *et al.* [@triki_Biological_2019;@triki_Brain_2020] 
tested the fish for a maximum of 200 trials with 20 trials a day, 10 trials 
in the morning and 10 trials in the afternoon. They randomized and 
counterbalanced the plates’ spatial location (i.e. left or right) 
between trials. Similarly, they counterbalanced the plates’ decoration 
(colour and pattern) and the plates’ role (visitor or resident) between the 
tested fish. In the original studies, once a fish reach a learning criterion, 
that is, performing significantly above chance level ($> 50%$, $p-value ≥ 0.05$),
they passed to a reversal version of the task where the roles of the 
visitor/resident Plexiglas plates were swapped (see Triki et al. 2019, 2020). 
Here, we used instead a subset of these data in order to have an idea on 
cleaner fish preferences for the visitor plate, even if they do not reach 
the learning criterion within 200 trials. To do so, we first extracted the 
trial-by-trial outcomes from the last two session (20 trials) of those who 
never reached the learning criterion for visitor plate 
($N = `r fieldData.cleaner.filt.sum[V1==2,length(V1)]`$ cleaner fish). 
For those who reached the learning criterion at some point during the 
test and passed to a reversal phase, we extracted the trials-by-trial 
outcomes from the last session (10 trials) before passing to reversal 
and the last 10 trials they were exposed to in the test 
($N = `r fieldData.cleaner.filt.sum[V1>2,length(V1)]`$ cleaner fish).

@@

(ref:post) Posterior distributions for parameter values $\gamma$ and $\eta$ for 
the three model ("Full model", "chaining only" and "penalty only"). We show 
the kernel density estimates, below the mode (black dot) and the 65% 
(light blue shade) and 95% (grey shade)  highest posterior density interval 
for the two parameters. On the top, panels a and b show posterior 
distributions for a full model, including chaining ($\gamma$) and penalty 
($\eta$). Panel c, shows the $\gamma$ estimate from model with only chaining. 
Panel d shows the $\\eta$ estimate from a model with only penalty. Panel e, 
shows a measure of fit for all models, namely the distribution of pseudo-$R^2$ 
obtained from sampling the posterior distribution of parameter values.

@@

# Name of the folders containing the MCMC runs
scen1<-"MCMCclean_gam_nRew_sca_"
scen2<-"MCMCclean_gam_sca_"
scen3<-"MCMCclean_Nrew_sca_"

@@

MCMCdata1<-loadMCMCrun(scen1,thinning = thinning,burn.in = burn.in)
MCMCdata2<-loadMCMCrun(scen2,thinning = thinning,burn.in = burn.in)
# MCMCdata3<-loadMCMCrun(scen3,thinning = thinning,burn.in = burn.in)

@@

# mcmcList.3<-mcmc.list(do.call(list,lapply(unique(MCMCdata3$seed), function(repli){
#   rundata<-MCMCdata3[seed==repli]
#   mcmcRun<-mcmc(rundata[,.(gamma,scaleConst)],thin = thinning)#
#   return(mcmcRun)
# })))

@@

# cuts.3<-lapply(MCMCdata3[,.(negReward,scaleConst)],
#                function(x){
#                  c(mode_hdci(x,.width = intervals[c(2,1)])$ymin,
#                    mode_hdci(x,.width = intervals)$ymax)
#                })

# get all the likelihoods from the posterior for all models
# loglikehoods.all<-data.table(lglikelihood=c(MCMCdata1$fit,MCMCdata2$fit,
#                                              MCMCdata3$fit),
#                               model=c(rep(labels.Scen[1],length(MCMCdata1$fit)),
#                                       rep(labels.Scen[2],length(MCMCdata2$fit)),
#                                       rep(labels.Scen[3],length(MCMCdata3$fit))
#                                       ))
# 
# # Read field data to calculate null log-likelihood
# fieldData<-fread(here("Data","data_cleaner_abs_threa1.5.txt"))
# 
# nullLikehood<-sum(dbinom(x=fieldData[,score_visitor],
#                          size = 20,prob = 0.5,log = TRUE))
# 
# # calculate pseudo R^2
# loglikehoods.all[,Rsqrd:=1-lglikelihood/nullLikehood]
# 
# 
# panel.rsqr.all<-
#   ggplot(data=loglikehoods.all[is.finite(lglikelihood)],
#                   aes(y=model,x=Rsqrd,fill=model))+
#   stat_halfeye(alpha=0.5,point_size=3)+
#   theme_classic()+#xlim(-0.5,0.25)+
#   scale_fill_manual(values=c("#1b9e77","#d95f02","#7570b3"),
#   name = "Model", labels = labelsPlot.Scen)+
#   scale_y_discrete(labels=labelsPlot.Scen)+
#   geom_vline(xintercept = 0,color="black",size=0.5)+
#   theme(legend.position = 'none')
#           # c(.32, .65)
#         # ,legend.key.size = unit(0.22,'cm'),
#         # axis.text.y = element_blank(),
#         # legend.text = element_text(size=7))
# 
# gam.both.post<-ggplot(data=MCMCdata1,aes(x=gamma)) +
#   stat_halfeye(aes(fill=stat(cut(x,breaks = cuts.1$gamma))),
#                point_interval = mode_hdi, .width = c(.66, .95),
#                point_size=3,show.legend=FALSE) +
#   labs(title=labelsPlot.Scen[1],
#        subtitle = expression(gamma),x="",y="")+
#   theme_classic()+
#   scale_fill_manual(values=c("gray85","skyblue","gray85"))
# 
# nrew.both.post<-ggplot(data=MCMCdata1,aes(x=negReward)) +
#   stat_halfeye(aes(fill=stat(cut(x,breaks = cuts.1$negReward))),
#                point_interval = mode_hdi,
#                .width = c(.66, .95),point_size=3,
#                show.legend=FALSE)+theme_classic()+
#   scale_fill_manual(values=c("gray85","skyblue","gray85"))+
#   labs(title="",subtitle = expression(eta),x="",y="")+
#   xlim(-2,2)
# gam.gam.post<-ggplot(data=MCMCdata2,aes(x=gamma)) +
#   stat_halfeye(aes(fill=stat(cut(x,breaks = cuts.2$gamma))),
#                point_interval = mode_hdi, .width = c(.66, .95),
#                point_size=3,
#                show.legend=FALSE) +
#   scale_fill_manual(values=c("gray85","skyblue","gray85"))+
#   labs(title=labelsPlot.Scen[2],
#        subtitle = expression(gamma),x="",y="")+
#   theme_classic()
# nrew.nrew.post<-ggplot(data=MCMCdata3,aes(x=negReward)) +
#   stat_halfeye(aes(fill=stat(cut(x,breaks = cuts.3$negReward))),
#                point_interval = mode_hdi, .width = c(.66, .95),
#                point_size=3,
#                show.legend=FALSE) +
#   scale_fill_manual(values=c("gray85","skyblue","gray85"))+
#   labs(title=labelsPlot.Scen[3],
#        subtitle = expression(gamma),x="",y="")+
#   theme_classic()+xlim(-2,2)
# 
# first.col<-plot_grid(nrow=3,align = "v",byrow = TRUE,
#           gam.both.post,gam.gam.post,nrew.nrew.post,
#           labels=c('a','c','d'))
# second.col<-plot_grid(nrow = 2,align = "v",byrow = TRUE,
#           nrew.both.post,panel.rsqr.all,rel_heights = c(1,2),
#           labels = c('b','e'))
# plot_grid(ncol = 2,first.col,second.col)

@@

<!-- # ```{r pred, fig.cap='(ref:pred)',out.width='100%',fig.height=5} -->
<!-- #  -->
<!-- # source(here("manFigures.R")) -->
<!-- # ggarrange(cont.obs.pred.both,scatter.obs.pred.both, -->
<!-- # cont.obs.pred.gam,scatter.obs.pred.gam, -->
<!-- # cont.obs.pred.Nrew,scatter.obs.pred.Nrew, -->
<!-- #           labels=c('a','b','c','d'),common.legend = F, -->
<!-- #   heights=c(1.25,1)) -->
<!-- #  -->
<!-- # ``` -->

@@

In this study, our main aim was to unravel which of two potential 
cognitive mechanisms, chaining of events, operant conditioning, 
or their combination, best explains wild-caught cleaner fish 
performance in the ephemeral reward task, while accounting for their
ecological conditions. To evaluate the merits of each of these two mechanisms
separately and combined, we considered cleaner fish performance 
in the lab test to have its origin from the these fish apply 
in their natural environment. That is, individuals that 
solved the task already had a preference for 
visitors clients, and generalized this rule to the lab conditions 
once being familiar with the task.

@@

```{r diagnosticsfull, fig.cap='(ref:diagnosticsfull)',echo=FALSE,fig.height=5}

@@

pdf(file = NULL)
gp.dat<-gelman.plot(mcmcList.1)
dev.off()

@@

pdf(file = NULL)
gp.dat.gam<-gelman.plot(mcmcList.2)
dev.off()

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,warning = FALSE,message=FALSE,cache = TRUE)
library(here)
source(file=here("loadFieldData.R"))
```

Often alternative cognitive mechanisms yield similar behaviour and/or cognitive
performances. This poses a problem for disentangling the mechanistic 
underpinnings of behaviour. This is particularly clear in research aimed at
discovering interspecific variation in *higher* cognitive abilities; or 
in other words, research on whether non-human animals show cognitive 
abilities believed to be uniquely human. For instance, when researchers 
try to find *mental time travel-like* behaviour, they usually come up with 
clever experiments showing the behaviour displayed requires inferences made 
through past events[@dally_Foodcaching_2006]. However, they often face challenges proposing alternative scenarios where simple associations can bring about the 
predicted behavioural outcome [@suddendorf_Evolution_2007]. Similarly, attempts 
to show *theory of mind* in non-human animals face objections 
justified by alternative mechanisms that bring about similar behavioural 
results [@heyes_Theory_1998]. Such controversies are usually settled by 
using the principle of parsimony, and it's cognitive version, 
Lloyd Morgan’s cannon. Which states that the simpler explanation (mechanism)
must be accepted. Ideally, however, alternative hypothesis should be evaluated 
in light of their explanatory power. One potential dimension that could allow
the variation necessary among distinct cognitive mechanisms is environmental
variation. Controlled laboratory conditions often mask the inter and 
intra-specific variation that arises in the interaction between mechanisms and 
environments. Thus, if predictions from the alternative mechanisms change
with environmental conditions, it is possible to asses the explanatory power 
of the different hypothesis by performing behavioural test in the context of
different environmental conditions.

From a evolutionary perspective, mechanisms are likely selected because 
of how they allow individuals to respond to environmental variation. 
For example, biological market theory predicts that the exchange rate of goods 
and/or services traded between cooperative partners adjusts to the 
law of supply and demand, when individuals have some degree of partner choice
[@noe_Biological_1995]. Supply and demand conditions, which typically 
depend on the abundance of the species involved, certainly vary in time 
and space. Therefore, natural selection should favour the ability 
to flexibly adjust decisions and behavioural output to 
current market conditions. Indeed, such adjustments have been repeatedly 
documented [@axen_Signalling_1996]. In animals, an obvious general candidate 
mechanism for the strategic adjustment is the cognitive machinery. However, 
it is not clear which cognitive mechanisms allow individuals to adjust their 
behaviour to the varying conditions. As adjustments to changes in 
market conditions have been documented in a great variety of species with 
highly variable brain anatomies, the question arises to what extent 
mechanisms beyond basic associative learning may be involved.

One example of strategic adjustment in a biological market is the marine
cleaning mutualism involving the cleaner fish *Labroides dimidiatus* and
'client' fish. Client fish seek cleaner fish services at their territory 
(so-called “cleaning station”) and offer themselves as food patches 
to get their ectoparasites removed, which provides cleaners 
with food and clients with improved health [@waldie_LongTerm_2011;
@ros_Does_2011;@triki_Effects_2016;@demaire_Reduced_2020]. 
Given the capacity of some client fish to swim larger distances and 
access multiple cleaning stations while others access the only cleaning 
station at their territory, it is crucial to categorize clients as 
either “visitors” or “residents”, respectively. During cleaning interactions, 
a cleaner fish often face a choice between a visitor and a resident client 
seeking its cleaning services simultaneously. Visitors have the option 
to switch to another cleaner fish if being made to wait, while residents 
must wait for inspection. Indeed,
visitors have been observed to use their partner choice option in that
way [@bshary_Choosy_2002], which may explain why cleaners give
visitors service priority in a field study in the Red Sea
[@bshary_Cleaner_2001].

When researchers aimed at testing wild-caught cleaner fish, as 
well as individuals from other species, abilities 
to prioritize visitor over resident clients in a lab-based paradigm, 
they used Plexiglas plates of different colours and shapes offering same 
amount of food as surrogates for visitor and resident clients.
One plate is made to behave like a resident, i.e. it remains until 
the cleaner fish had eaten off it. The other plate was made to behave 
like a visitor, i.e. it is removed if not inspected first. 
This laboratory paradigm originally inspired by the cleaning market 
of cleaner fish is more generally refered as the ephemeral reward task. 
Individuals are assesed in their capacity to develop a preference for the 
ephemeral option, which gives them access to two reward units.
Cleaners learned to prefer the visitor plate and
hence obtained the double amount of food [@bshary_Asymmetric_2002].
Furthermore, adult cleaner fish outperformed various primates as well as rats and
pigeons in this original version of the ephemeral reward task
[@salwiczek_Adult_2012; @zentall_Early_2017]; yet the African grey
parrots solve this task as well [@pepperberg_Can_2014]. This interspecific
variation in performance, and primates' known cognitive abilities $\color{red}{\text{ref}}$, shows that performane in the task
cannot be readily explained by classifying species as having high and 
low cognitive abilities. Instead, this inter-spefic variation suggest
that solving the ephemeral reward task requires a set of cognitive 
mechanisms that are not present in all the tested taxa, or were not used 
under the experimental set-up.

@@

To solve the ephemeral reward task, cleaners need to account for 
the future consequences of current decisions. In the model by 
Quiñones *et al.* [@quinones_Reinforcement_2019], this could be achieved in 
two non-mutually exclusive ways: through low temporal discounting of 
future effects, also termed ‘chaining’ [@enquist_Power_2016]; and/or 
through perceiving a visitor client leaving as psychological punishment 
(i.e. as a negative reinforcer). Low temporal discounting is when 
individuals include in their valuation of an action the reward effects 
that this will have in the future. This is done by combining in a single 
valuation the reward obtained in the current time with all the reward that 
comes after, discounting for how far in the future reward 
is accrued. 'Chaining' the reward of these different time steps allows 
individuals to take actions that increase the long-term reward at the sacrifice
of short term considerations [@enquist_Power_2016]. Even though, 'chaining' 
can be readily implemented computationally in learning models 
[@enquist_Power_2016;@sutton_Reinforcement_2018], cognitively it seems to be 
a complex adaptation [@suddendorf_Evolution_2007]. On the other hand,
using client behaviour as a negative reinforcer is in principle 
easier to implement. Negative reinforcement is part of the ubiquitous 
learning mechanism termed operant conditioning [@thorndike_Animal_1898; 
@skinner_Behavior_1938]. Thus, standard logic of 
Lloyd Morgan’s cannon demands that operant conditioning as the simpler 
explanation is to be accepted by default. Ideally, however, the two 
mechanisms should be evaluated in light of how well 
they explain the available data. Note that different fields interested in
cognition and desicion making use different words to refer to negative 
reinforcers [@quinones_Reinforcement_2019;@sutton_Reinforcement_2018]. 
In here, for the sake of simplicity and clarity, we will use the word 
'penalty' to refer to this mechanism which includes a negative reinforcer.

@@

The interplay between the underling cognitive machinery and 
local ecological conditions apparently generate the documented variable 
performance among individuals of the same species. 
Our approach of fitting the computational model to the empirical data on 
fish densities and cleaner fish censuses and performance
in the ephemeral reward task aimed at: (i), determining which 
mechanism cleaner fish use to incorporate future consequences of current 
decisions by testing whether chaining, penalty, or a combination 
of both best explains their performance; (ii)  determining whether the 
two mechanisms differed with respect to the ecological conditions 
that are likely to cause high versus low performance in the 
ephemeral reward task. Additionally, we assessed which mechanism 
yields optimal performance patterns. Relying on the logic of 
biological market theory, we predicted 
that appropriate performance is to show low preference for visitors under low
local cleaner-to-client ratio.

@@

The model shows that agents need to find a mean 
to incorporate future consequences of current choices. In the model, 
this could be achieved with either of two parameters that could 
also work together.  First, $\gamma$ measures how
much individuals include future rewards in their decision updates. 
If $\gamma=0$, individuals only use the immediate reward obtained from a
cleaning interaction. As $\gamma$ increases individuals include more the
reward obtained from the subsequent choices. That amounts to calculating
and using for decision making the future expected rewards of an action (chaining).
Second, $\eta$ measures how much individuals include in their reward the
fleeing behaviour of visitor as a negative component (penalty). Both of
these parameters allow individuals to use in their estimates the future
effects of their choices.

@@

@@

The cognitive performance data was from a total of 120 cleaners 
(10 individuals per 12 site/year) tested in the 
ephemeral reward task [@triki_Biological_2019;@triki_Brain_2020]. Authors 
housed all captured cleaners individually in glass aquaria 
( $62cm \times 27cm \times 37 cm$ ) and provided them 
with PVC pipes ($10 cm \times 1 cm$) as shelters. 
The task consisted of exposing the cleaner fish to substitute 
models of client fish in the form of two *Plexiglas* plates offering the 
same amount of food (one item of mashed prawn).  The two plates differed 
in colour and pattern (horizontal green stripes or vertical pink stripes) 
but had equal size ($10 cm \times 7 cm$). Importantly, the two plates played 
different roles as either a visitor (ephemeral food source) or 
resident (permanent food source). That is, if a cleaner fish inspected the 
resident plate first, the experimenter withdrew the visitor plate out of 
the aquarium as a consequence.  Choosing first the visitor plate, 
however, granted access to both plates. The equal size of the plates 
forced cleaner fish to learn to give service priority to the visitor plate 
based solely on the behaviour-cue of the plates rather than size-cue 
[@wismer_Cuebased_2019]. Triki *et al.* [@triki_Biological_2019;@triki_Brain_2020] 
tested the fish for a maximum of 200 trials with 20 trials a day, 10 trials 
in the morning and 10 trials in the afternoon. They randomized and 
counterbalanced the plates’ spatial location (i.e. left or right) 
between trials. Similarly, they counterbalanced the plates’ decoration 
(colour and pattern) and the plates’ role (visitor or resident) between the 
tested fish. In the original studies, once a fish reach a learning criterion, 
that is, performing significantly above chance level ($> 50%$, $p-value ≥ 0.05$),
they passed to a reversal version of the task where the roles of the 
visitor/resident Plexiglas plates were swapped (see Triki et al. 2019, 2020). 
Here, we used instead a subset of these data in order to have an idea on 
cleaner fish preferences for the visitor plate, even if they do not reach 
the learning criterion within 200 trials. To do so, we first extracted the 
trial-by-trial outcomes from the last two session (20 trials) of those who 
never reached the learning criterion for visitor plate 
($N = `r fieldData.cleaner.filt.sum[V1==2,length(V1)]`$ cleaner fish). 
For those who reached the learning criterion at some point during the 
test and passed to a reversal phase, we extracted the trials-by-trial 
outcomes from the last session (10 trials) before passing to reversal 
and the last 10 trials they were exposed to in the test 
($N = `r fieldData.cleaner.filt.sum[V1>2,length(V1)]`$ cleaner fish). 
We chose a combination of initial and reversal to quantify preference 
for the visitor client, given that it matches well the criteria chosen 
in previous analysis of the ephemeral reward task experimental set-up
[@triki_Biological_2019;@triki_Brain_2020]. In the 
suplementaty material (Fig. \ref{fig:rawdata}) we show how our choice maps
to the previously used criteria. For comparison, we also show a quantification 
of visitor preference based only on the initial phase of the experiments.

@@

(ref:post) Posterior distributions for parameter values $\gamma$ and $\eta$ for 
the three models ("Full model", "chaining only" and "penalty only"). We show 
the kernel density estimates, below the mode (black dot) and the 65% 
(light blue shade) and 95% (grey shade)  highest posterior density interval 
for the two parameters. On the top, panels a and b show posterior 
distributions for a full model, including chaining ($\gamma$) and penalty 
($\eta$). Panel c, shows the $\gamma$ estimate from model with only chaining. 
Panel d shows the $\\eta$ estimate from a model with only penalty. Panel e, 
shows a measure of fit for all models, namely the distribution of pseudo-$R^2$ 
obtained from sampling the posterior distribution of parameter values.

@@

# Name of the folders containing the MCMC runs
# scen1<-"MCMCclean_gam_nRew_sca_"
# scen2<-"MCMCclean_gam_sca_"
# scen3<-"MCMCclean_Nrew_sca_"

scen1<-"MCMCclean_gam_Nrew_sca_old_"
scen2<-"MCMCclean_gam_sca_old_"
scen3<-"MCMCclean_Nrew_sca_old_"

@@

MCMCdata1<-loadMCMCrun(scen1,thinning = thinning,burn.in = burn.in)
MCMCdata2<-loadMCMCrun(scen2,thinning = thinning,burn.in = burn.in)
MCMCdata3<-loadMCMCrun(scen3,thinning = thinning,burn.in = burn.in)

@@

mcmcList.3<-mcmc.list(do.call(list,lapply(unique(MCMCdata3$seed), function(repli){
  rundata<-MCMCdata3[seed==repli]
  mcmcRun<-mcmc(rundata[,.(negReward,scaleConst)],thin = thinning)#
  return(mcmcRun)
})))

@@

cuts.3<-lapply(MCMCdata3[,.(negReward,scaleConst)],
               function(x){
                 c(mode_hdci(x,.width = intervals[c(2,1)])$ymin,
                   mode_hdci(x,.width = intervals)$ymax)
               })

# get all the likelihoods from the posterior for all models
loglikehoods.all<-data.table(lglikelihood=c(MCMCdata1$fit,MCMCdata2$fit,
                                             MCMCdata3$fit),
                              model=c(rep(labels.Scen[1],length(MCMCdata1$fit)),
                                      rep(labels.Scen[2],length(MCMCdata2$fit)),
                                      rep(labels.Scen[3],length(MCMCdata3$fit))
                                      ))

# Read field data to calculate null log-likelihood
fieldData<-fread(here("Data","data_cleaner_abs_threa1.5.txt"))

nullLikehood<-sum(dbinom(x=fieldData[,score_visitor],
                         size = 20,prob = 0.5,log = TRUE))

# calculate pseudo R^2
loglikehoods.all[,Rsqrd:=1-lglikelihood/nullLikehood]

panel.rsqr.all<-
  ggplot(data=loglikehoods.all[is.finite(lglikelihood)],
                  aes(y=model,x=Rsqrd,fill=model))+
  stat_halfeye(alpha=0.5,point_size=3)+
  theme_classic()+xlim(-0.5,0.25)+
  scale_fill_manual(values=c("#1b9e77","#d95f02","#7570b3"),
  name = "Model", labels = labelsPlot.Scen)+
  scale_y_discrete(labels=labelsPlot.Scen)+
  geom_vline(xintercept = 0,color="black",size=0.5)+
  theme(legend.position = 'none')
          # c(.32, .65)
        # ,legend.key.size = unit(0.22,'cm'),
        # axis.text.y = element_blank(),
        # legend.text = element_text(size=7))

gam.both.post<-ggplot(data=MCMCdata1,aes(x=gamma)) +
  stat_halfeye(aes(fill=stat(cut(x,breaks = cuts.1$gamma))),
               point_interval = mode_hdi, .width = c(.66, .95),
               point_size=3,show.legend=FALSE) +
  labs(title=labelsPlot.Scen[1],
       subtitle = expression(gamma),x="",y="")+
  theme_classic()+
  scale_fill_manual(values=c("gray85","skyblue","gray85"))

nrew.both.post<-ggplot(data=MCMCdata1,aes(x=negReward)) +
  stat_halfeye(aes(fill=stat(cut(x,breaks = cuts.1$negReward))),
               point_interval = mode_hdi,
               .width = c(.66, .95),point_size=3,
               show.legend=FALSE)+theme_classic()+
  scale_fill_manual(values=c("gray85","skyblue","gray85"))+
  labs(title="",subtitle = expression(eta),x="",y="")+
  xlim(-2,2)
gam.gam.post<-ggplot(data=MCMCdata2,aes(x=gamma)) +
  stat_halfeye(aes(fill=stat(cut(x,breaks = cuts.2$gamma))),
               point_interval = mode_hdi, .width = c(.66, .95),
               point_size=3,
               show.legend=FALSE) +
  scale_fill_manual(values=c("gray85","skyblue","gray85"))+
  labs(title=labelsPlot.Scen[2],
       subtitle = expression(gamma),x="",y="")+
  theme_classic()
nrew.nrew.post<-ggplot(data=MCMCdata3,aes(x=negReward)) +
  stat_halfeye(aes(fill=stat(cut(x,breaks = cuts.3$negReward))),
               point_interval = mode_hdi, .width = c(.66, .95),
               point_size=3,
               show.legend=FALSE) +
  scale_fill_manual(values=c("gray85","skyblue","gray85"))+
  labs(title=labelsPlot.Scen[3],
       subtitle = expression(gamma),x="",y="")+
  theme_classic()+xlim(-2,2)

first.col<-plot_grid(nrow=3,align = "v",byrow = TRUE,
          gam.both.post,gam.gam.post,nrew.nrew.post,
          labels=c('a','c','d'))
second.col<-plot_grid(nrow = 2,align = "v",byrow = TRUE,
          nrew.both.post,panel.rsqr.all,rel_heights = c(1,2),
          labels = c('b','e'))
plot_grid(ncol = 2,first.col,second.col)

@@

```{r pred, fig.cap='(ref:pred)',out.width='100%',fig.height=5}

 source(here("manFigures.R"))
 ggarrange(cont.obs.pred.both,scatter.obs.pred.both,
 cont.obs.pred.gam,scatter.obs.pred.gam,
 cont.obs.pred.Nrew,scatter.obs.pred.Nrew,
           labels=c('a','b','c','d','e','f'),common.legend = F,
   heights=c(1.4,1,1),nrow = 3,ncol = 2)

@@

In this study, our main aim was to unravel which of two potential 
cognitive mechanisms, chaining of events, penalty, 
or their combination, best explains wild-caught cleaner fish 
performance in the ephemeral reward task, while accounting for their
ecological conditions. To evaluate the merits of each of these two mechanisms
separately and combined, we considered cleaner fish performance 
in the lab test to have its origin from the these fish apply 
in their natural environment. That is, individuals that 
solved the task already had a preference for 
visitors clients, and generalized this rule to the lab conditions 
once being familiar with the task.

@@

```{r diagnosticsfull, fig.cap='(ref:diagnosticsfull)',message=FALSE,echo=FALSE,fig.height=5}

@@

pdf(file = NULL)
gp.dat<-gelman.plot(mcmcList.1)
rubbish <- dev.off()

@@

pdf(file = NULL)
gp.dat.gam<-gelman.plot(mcmcList.2)
rubbish <- dev.off()
